---
title: "Examining site effects in gadgetron volumes (v3)"
author: "Virgilio Gonzenbach"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    toc: True
    toc_float: True
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(kableExtra)

df = read.csv('data/gadgetron_volumes_v3.3.csv', stringsAsFactors = TRUE, row.names = NULL)
df$source_path = NULL
df$X = NULL
vol.names = df %>% select(where(is.numeric)) %>% colnames
#qc_df = readxl::read_xlsx('data/mprage_skull_strip_QC.xlsx', sheet=1)
#qc_df = qc_df[rowSums(is.na.data.frame(qc_df)) <= 2,] # Exclude bottom rows
```

## Data summary

### Raw Data Tally

We are missing 3 raw data files in total that we **should** have:

```{r}
raw_tally_df = read.csv('tally_raw_data.csv')
raw_tally_df
```

Note: subject 03-001 was not scanned at NIH site; subject 03-002 was not scanned at Penn.

### Total visits per subjects

```{r}
df %>% count(ID, img_type) # %>% select(-img_source, -source_path) %>% count()
```

### Scans at site per subject

```{r}
subj_df_list = split(df, df$ID)
subj_df_list %>% sapply(function(df) table(df[, "site"])) %>% t() %>% as.data.frame()
```

### Missing segmentations

Showing only segmentation types with at least 1 file missing.

```{r}
miss_vol = as.data.frame(is.na.data.frame(select(df, all_of(vol.names)))) 
row_filter = rowSums(miss_vol) > 0 # Sum TRUE values
col_filter = names(which(colSums(miss_vol) > 0)) 

# Transform to character values
miss_vol = miss_vol %>%  lapply(function(x) as.character(x) %>% dplyr::recode(`TRUE` = 'MISSING', `FALSE` = '')) %>% bind_cols
miss_df = cbind(select(df, -all_of(vol.names)), miss_vol) %>% filter(row_filter) %>% select(-all_of(setdiff(vol.names, col_filter)))

miss_df
```

## Segmentation volumetrics

```{r}
# Load onscanner data
onsc_df = read.csv('data/onscanner_volumes.csv')
```

```{r, warning=FALSE}
local.vols = c('mimosa', 'JLF_thal', 'FIRST_thal')
global.vols = setdiff(vol.names, local.vols)
  

plot.vols = function(df, filt, color='ID'){
  df %>% pivot_longer(cols=vol.names, names_to = "segmentation", values_to = "cm") %>% filter(segmentation %in% filt) %>% ggplot(aes(x=segmentation, y=cm, color = ID)) + geom_jitter() + facet_grid(~site) + coord_flip()
  
}  

p = plot.vols(df, global.vols)
p + ggtitle("Global Volumes (Gadgetron)") + theme(plot.title = element_text(hjust = 0.5)) + ylim(0,1200)
```

```{r,  warning=FALSE}
p = plot.vols(onsc_df, global.vols)
p + ggtitle("Global Volumes (Onscanner)") + theme(plot.title = element_text(hjust = 0.5)) + ylim(0,1200)
```

```{r, warning=FALSE}
p = plot.vols(df, local.vols)
p + ggtitle("Local Volumes (Gadgetron)") + theme(plot.title = element_text(hjust = 0.5))+ ylim(0, 80)
```

```{r, warning=FALSE}
p = plot.vols(onsc_df, local.vols)
p + ggtitle("Local Volumes (Onscanner)") + theme(plot.title = element_text(hjust = 0.5)) + ylim(0, 80)
```

## Site effects

```{r, warning=FALSE}
get.r.sq = function(df){

 r.sq = numeric(0)
 for (v in vol.names) { # global vol.names
   # deal with NAs
   res = tryCatch(summary(lm(df[,v ] ~ df$site))$r.squared, error = function(cond) return(NA))
   r.sq = c(r.sq, res)
 }
 names(r.sq) = vol.names
 return(r.sq)
}

# Get r-sq for gadgetron recons
R.sq.df = subj_df_list %>% 
  lapply(get.r.sq) %>% 
  bind_rows(.id='ID') %>% 
  tibble::add_column(recon = 'gadgetron', .before = "ID")

R.sq.df = split(onsc_df, f= onsc_df$ID) %>% 
  lapply(get.r.sq) %>% 
  bind_rows(.id='ID') %>% 
  tibble::add_column(recon = 'onscanner', .before = "ID") %>% 
  rbind(R.sq.df)

R.sq.df %>% 
  pivot_longer(cols=vol.names, names_to = "segmentation", values_to = "R.squared") %>% 
  ggplot(aes(x=segmentation, y=R.squared, color = recon)) + 
  geom_boxplot() + 
  coord_flip() + 
  ggtitle("R squared values of site effects") + 
  theme(plot.title = element_text(hjust = 0.5))
# R.sq.df %>% 
#   pivot_longer(cols=vol.names, names_to = "segmentation", values_to = "R.squared") %>% 
#   ggplot(aes(x=segmentation, y=R.squared, color = ID)) + 
#   geom_boxplot(aes(x=segmentation, y=R.squared, color = NULL)) +
#   geom_jitter() +
#   coord_flip() +
#   ggtitle("R squared values of site effects (Gadgetron)") + theme(plot.title = element_text(hjust = 0.5))
```

Paired t-tests didnot find a significant difference.

```{r}
vol.names %>% lapply(function(x) t.test(reformulate("recon", x), data = subset(R.sq.df, recon %in% c("onscanner", "gadgetron")), paired = TRUE))
```

## Permutation Testing

### Gadgetron Volumes

pseudo-F Ratio statistics: Intersite Variability / Intrasite Variability

```{r}
gadg_ratio = readRDS("results/gadgetron_volumes_v3.3_ratio_stat.rds")
gadg_null = readRDS("results/gadgetron_volumes_v3.3_null.rds")
gadg_ratio$X = NULL
gadg_ratio
```

Distribution of permutated pseudo-F stats:

```{r, warning=FALSE}
gadg_null %>% pivot_longer(everything()) %>% ggplot(aes(x=name, y=value)) + geom_violin(draw_quantiles = c(0.95)) + geom_point(aes(x=name, y=value), data = gadg_ratio %>% pivot_longer(everything())) + coord_flip()
```

P-value of pseudo-F stat:

```{r}
test_ratio_stat = function(ratio.stats, null.dists){
  p.vals = c()
  for (col in colnames(ratio.stats)){
    ratio.stat = ratio.stats[, col]
    null.dist = null.dists[, col]
    
    percentile = ecdf(null.dist)
    p.vals = c(p.vals, percentile(ratio.stat))
  }
  names(p.vals) = colnames(ratio.stats)
  return(p.vals)
}

p = 1 - test_ratio_stat(gadg_ratio, gadg_null)
(10000*p + 1)/(1+10000)
```

### Onscanner Volumes

pseudo-F statistics: Intersite Variability / Intrasite Variability

```{r}
onsc_ratio = readRDS("results/onscanner_volumes_ratio_stat.rds")
onsc_null = readRDS("results/onscanner_volumes_null.rds")

onsc_ratio
```

Distribution of permutated pseudo-F stats:

```{r, warning=FALSE}
onsc_null %>% pivot_longer(everything()) %>% ggplot(aes(x=name, y=value)) + geom_violin(draw_quantiles = c(0.95)) + geom_point(aes(x=name, y=value), data = pivot_longer(onsc_ratio, everything())) + coord_flip()
```

P-value of pseudo-F stat:

```{r}
p = 1 - test_ratio_stat(onsc_ratio, onsc_null)
(10000*p + 1)/(1+10000)
```

## Mixed Models

```{r, warning=FALSE, message=FALSE}
library(lme4)
library(dplyr)
library(sjPlot)
library(jtools)
library(rlang)
library(EMAtools)
library(car)

#Mixed Effects Model, random effect term for subject and another random effect term that nests site within subject
gadg_df = df

run_mixed_models = function(vol_df){
  
  vols = vol_df %>% select(where(is.numeric)) %>% colnames
  models = list()
  for (vol in vols){ 
    x = vol_df[, vol]
    assign(vol, x)
    message(sprintf("Ran mixed model for %s", vol))
    model = lmer(reformulate(" (1|ID) + (1|site:ID)", vol), data = vol_df)
    models = c(models, model)
  }
  names(models) = vols 
  return(models)
}

gadg_models = run_mixed_models(gadg_df)
gadg_ICC_df = gadg_models %>% lapply(performance::icc, by_group = TRUE) %>% bind_rows(.id = "volume") %>% pivot_longer(everything(), names_to = "volume", values_to = "ICC")
gadg_ICC_df = data.frame(volume = gadg_ICC_df$volume, Group = gadg_ICC_df$ICC$Group, ICC = gadg_ICC_df$ICC$ICC) %>% drop_na()

onsc_models = run_mixed_models(onsc_df)
onsc_ICC_df = onsc_models %>% lapply(performance::icc, by_group = TRUE) %>% bind_rows(.id = "volume") %>% as.data.frame()
#onsc_ICC_df = data.frame(volume = onsc_ICC_df$name, Group = onsc_ICC_df$value$Group, value = onsc_ICC_df$value$ICC) %>% drop_na()
```

```{r}
gadg_models %>% lapply(summary)
```

### ICC of Subject Random Effect

```{r, warning=FALSE}
gadg_ICC_df %>% filter(Group == 'ID') %>% ggplot(aes(x=volume, y=ICC)) + geom_point() + coord_flip() + ggtitle("ICC from Gadgetron Volumes") + ylim(0,1)
onsc_ICC_df %>% filter(Group == 'ID') %>% ggplot(aes(x=volume, y=ICC)) + geom_point() + coord_flip() + ggtitle("ICC from Onscanner Volumes") + ylim(0,1)
```

### ICC of Site:Subject Random Effect

```{r, warning=FALSE}
gadg_ICC_df %>% filter(Group == 'site:ID') %>% ggplot(aes(x=volume, y=ICC)) + geom_point() + coord_flip() + ggtitle("ICC from Gadgetron Volumes") + ylim(0,1)

onsc_ICC_df %>% filter(Group == 'site:ID') %>% ggplot(aes(x=volume, y=ICC)) + geom_point() + coord_flip() + ggtitle("ICC from Onscanner Volumes") + ylim(0,1)
```

## Mixed model Diagnostics

### FAST TBV

```{r, warning=FALSE}
perf_mixed_model = function(model){
  
  list(summary = summary(model),
       perf = performance::icc(model, by_group = TRUE),
       shapiro.resid = shapiro.test(residuals(model)),
       qq.resid = car::qqPlot(residuals(model),dist="norm", main="Residuals"),
       shapiro.ID = shapiro.test(coef(model)$`ID`[,1]),
       `shapiro.site:ID` = shapiro.test(coef(model)$`site:ID`[,1]),
       
       qq.ranef.ID = car::qqPlot(ranef(model)$ID[,1],dist="norm", main="Rand. Eff. ID"),
       `qq.ranef.site:ID` = car::qqPlot(ranef(model)$`site:ID`[,1],dist="norm", main="Rand. Eff. Site:ID"),
       res.v.fit = ggplot(mapping = aes(y = resid(model), x = fitted(model))) + geom_abline(intercept = 0, slope = 0) + geom_point() + geom_smooth() + theme_bw()
       )
}

models = gadg_models
perf_mixed_model(models$FAST_TBV)
```

### JLF GM

```{r}
perf_mixed_model(models$JLF_GM)
```

### JLF WM

```{r}
perf_mixed_model(models$JLF_WM)
```

### FIRST thal

```{r}
perf_mixed_model(models$FIRST_thal)
```

### JLF thal

```{r}
perf_mixed_model(models$JLF_thal)
```

### Mimosa

```{r}
perf_mixed_model(models$mimosa)
```
